\large Neural and Symbolic Optimization of Cosmological Particle-Mesh Simulation \vspace{0.3em}

### 2024-02-28 docs/papers/sto/so.tex
WIP draft training NNs

In this work, we use the loss function in \eqref{loss_power_ln} for
both $\vs(\vq)$ and $\delta(\vx)$.
We combine the two fields to form the total loss for a single snapshot,
and further sum over all snapshots in a single forward simulation
%
\begin{equation}
\cJ = \sum_{a_i} \left[\cL_{\vs(a_i)} + \cL_{\delta(a_i)}\right].
\end{equation}
%

@@

We perform distributed data-parallel training using GPU devices hosted
on multiple CPU nodes.
With each process being binded to a single GPU device, the training data
samples are uniformly distributed to the processes.
In total, we use [...] \GPU\ GPUs, each accompanied with [...] CPU cores
and [...] GB CPU memory mainly for preloading the training data.
In each training step, a data sample is fed into the GPU from the CPU
memory where the preloading takes place to save the reading time from
the disk.

For each training step, we configure and run \pmwd\ with the same
initial coniditions, cosmological parameters and other common
configurations as the training data sample.
As discussed in \autoref{sec:loss}, the loss in a single simulation sums
over all snapshots.
Our batch size is [...], i.e. the number of GPU devices, since one
simulation is run on a single GPU and compared to the corresponding
training data sample for loss.
Then the loss and gradients are collected and averaged across all GPU
devices, with which the optimizer is also updated.
The SO parameters are initialized to be the same for all devices, and
hence remain the same for all processes given the same averaged
gradients.
We use the Adam \citep{Adam} optimizer.

The SO NN parameters are initialized in a way such that the output
is very close to one, i.e. no SO.
We achieve this by setting the kernel of the output layer to be very
small random values and the bias to be zero, which is around one after
the exponential function.

Besides the mesh for the force evaluation, we also need meshes when
computing the power spectra of the density field for the loss function
as well as for visualization.

Hyperparameters:
\begin{itemize}
  \item Model: number of hidden layers and nodes of f and g nets,
  activation function
  \item Optimization: learning rate scheduler, momentum, weight decay
  and dropout rate etc
  \item Loss: fields to include and loss function forms
\end{itemize}

We use Optuna \citep{optuna_2019} to search for the optimal
hyperparameters.

We empoly a learning rate scheduler that reduces the learning rate on
plateau where the loss function stops decreasing.
The initial learning rate is set at [...] with a reduction factor of
[...] after [...] epochs of no decreasing on loss.

We implement the neural networks with \texttt{Flax}
\citep{flax2020github}, and use the optimizers in \texttt{Optax}, part
of the DeepMind JAX Ecosystem
\citep{deepmind2020jax}.

>>>>

\YZ{In this work, we use the loss function in \eqref{loss_power_ln} for
both $\vs(\vq)$ and $\delta(\vx)$.}
We combine the two fields to form the total loss for a single snapshot,
and further sum over all snapshots in a single forward simulation
%
\begin{equation}
\cJ = \sum_{a_i} \left[\cL_{\vs(a_i)} + \cL_{\delta(a_i)}\right].
\end{equation}
%

@@

We perform distributed data-parallel training using GPU devices hosted
on multiple computing nodes.
Each process is binded to a single GPU device, and the training data
samples are uniformly distributed to the processes.
In total, we use XXX \GPU\ GPUs, each accompanied with XXX CPU cores
and XXX GB CPU memory mainly for preloading the training data.
The order of data samples within each process is shuffled across
training epochs.
In each training step, a data sample is fed into the GPU from the CPU
memory where the preloading took place to save the repeated reading time
from the disk.

For each training step, we configure and run \pmwd\ with the same
initial coniditions, cosmological parameters and other common
configurations as the training data sample.
As discussed in \autoref{sec:loss}, the loss in a single simulation sums
over all snapshots.
Our batch size is XXX, i.e. the number of GPU devices, since one
simulation is run on a single GPU and compared to the corresponding
training data sample for loss.
Then the loss and gradients are collected and averaged across all GPU
devices, with which the optimizer is updated.
We use the Adam \citep{Adam} optimizer.

We initialize the parameters of NNs in a way such that the output is
very close to one, i.e. no SO.
This is achieved by setting the kernel of the output layer to be very
small random values and the bias to be zero, whose return is then around
one after the exponential function.
The parameters are also initialized to be the same for all devices, and
hence remain the same during training given the same averaged gradients.

% We use Optuna \citep{optuna_2019} to search for the optimal
% hyperparameters.

% We empoly a learning rate scheduler that reduces the learning rate on
% plateau where the loss function stops decreasing.
% The initial learning rate is set at [...] with a reduction factor of
% [...] after [...] epochs of no decreasing on loss.

Besides the mesh for the force evaluation, we also need meshes when
estimating the density field for the loss function as well as for
visualization.

Same as the \pmwd\ library, this work is also implemented using JAX
\citep{JAX}.
We implement the NNs with \texttt{Flax} \citep{flax2020github}, and use
the optimizers provided in \texttt{Optax}, part of the DeepMind JAX
Ecosystem \citep{deepmind2020jax}.
